\begin{acknowledgements} 

I would like to acknowledge the Georgia Tech Research Institute for the Robert G.\ Shackelford Fellowship and the George W.\ Woodruff School of Mechanical Engineering for the Presidential Fellowship.

I am grateful to my advisors, Dr.\ Chris Paredis and Dr.\ Comas Haynes, for their support and patience.  I thank Dr.\ Robert Moore for providing the initial opportunity to pursue this line of research and for the wisdom and guidance he provided along the way.  I thank the rest of my committee, Dr.\ Thomas Fuller, Dr.\ Tequila Harris, and Dr.\ Sheldon Jeter, for their careful review and valuable comments.

I appreciate the caring attitude and professionalism of the staff of the George W.\ Woodruff School of Mechanical Engineering.  They have always been extremely helpful, positive, and understanding.

I am thankful for my labmates who have offered ideas, moral support, and friendship:
Bill Bailey,
Bill Binder,
Douglas Broadwell,
Chris Ford, 
Sebastian Herzig,
Katie Hornbostel,
Dimitri Hughes,
Alisha Kasam, 
Alek Kerzhner, 
Ben Lee, 
Malik Little,
Rich Malak,
Roxanne Moore, 
George Nelson,
Axel Reichwein, 
Muhammad Salman,
Brian Taylor, and
Stephanie Thompson.  
I appreciate the help of Kevin Bandy with data analysis.

I thank Mike Angelo and Guido Bender for providing the benchmark fuel cell data from the Hawaii Natural Energy Institute.  

I would also like to acknowledge the open-source software community.  Many free tools have been very helpful, including 
Modelica \cite{Modelica3.3} and the Modelica Standard Library \cite{ModelicaSL3.2}, 
Python,
SciPy and NumPy \cite{Oliphant2007},
matplotlib \cite{Hunter2007}, 
Wikipedia, 
Inkscape, 
LibreOffice, 
LaTeX,
Kile,
JabRef,
git, 
Linux, and 
Ubuntu.

Finally, I thank my family.  I have been blessed with a family who has given me confidence, encouragement, and endless support.  They are so caring and empathetic that they have unfortunately felt the burden of this work with me.  My dad patiently listened as I thought through many of the ideas in this dissertation.

\end{acknowledgements} 